---
title: "Cross_validation"
author: "Ke Xu"
date: "11/28/2021"
output: github_document
---

Cross validation provides a way to compare the predictive performance of competing methods
```{r}
library(tidyverse)
library(modelr)
library(mgcv)

set.seed(1)
```

```{r}
# Non-linear distribution
nonlin_df = 
  tibble(
    id = 1:100,
    x = runif(100, 0, 1),
    y = 1 - 10 * (x - .3) ^ 2 + rnorm(100, 0, .3)
  )

nonlin_df %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point()
```

Create splits(testing & training) by hand; plot; fit some models
```{r}
train_df = sample_n(nonlin_df, 80)
test_df = anti_join(nonlin_df, train_df, by = "id")
# anti_join function: find those data who are not overlap in these two data sets

ggplot(train_df, aes(x = x, y = y)) + 
  geom_point() + 
  geom_point(data = test_df, color = "red")
```

Fit my models:
```{r}
# different types of models
linear_mod = lm(y ~ x, data = train_df)
smooth_mod = mgcv::gam(y ~ s(x), data = train_df) # s(), a smooth function
wiggly_mod = mgcv::gam(y ~ s(x, k = 30), sp = 10e-6, data = train_df)

train_df %>% 
  add_predictions(smooth_mod) %>%  #calculate prediction by fitted model
  ggplot(aes(x = x, y = y)) + geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

```{r}
train_df %>% 
  add_predictions(wiggly_mod) %>% 
  ggplot(aes(x = x, y = y)) + geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

Quantify the results, compute root mean squared errors (RMSEs) for each model
```{r}
rmse(linear_mod, test_df) # use test data to find the differences between fitted value and predicted value, the smaller rmse is, the better model
```

## CV iteratively
use `r modelr::crossv_mc`
```{r}
cv_df = 
  crossv_mc(nonlin_df, 100) 

cv_df %>% pull(train) %>% .[[1]] %>% as_tibble

# OR, use 'map'
cv_df = 
  crossv_mc(nonlin_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

Let's fit some models
```{r}
cv_df = 
  cv_df %>% 
  mutate(
    linear_mod  = map(train, ~lm(y ~ x, data = .x)),
    smooth_mod  = map(train, ~mgcv::gam(y ~ s(x), data = .x)),
    wiggly_mod  = map(train, ~gam(y ~ s(x, k = 30), sp = 10e-6, data = .x))) %>% 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, ~rmse(model = .x, data = .y)),
    rmse_smooth = map2_dbl(smooth_mod, test, ~rmse(model = .x, data = .y)),
    rmse_wiggly = map2_dbl(wiggly_mod, test, ~rmse(model = .x, data = .y)))
```

Look at output
```{r}
# violin plot of rmse
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(), # means rmse_linear:rmse_wiggly
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

```{r}
# boxplot of rmse
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    rmse_linear:rmse_wiggly,
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_boxplot()
```


## Child growth

```{r}
child_growth = read_csv("./class2_data/nepalese_children.csv")

# The plots suggests some non-linearity, especially at the low end of the weight distribution. We’ll try three models: a linear fit; a piecewise linear fit; and a smooth fit using gam.
```

Consider models
```{r}
# For piecewise model
child_growth =
  child_growth %>% 
  mutate(weight_cp = (weight > 7) * (weight - 7))

linear_mod = lm(armc ~ weight, data = child_growth)
pwl_mod    = lm(armc ~ weight + weight_cp, data = child_growth)
smooth_mod = gam(armc ~ s(weight), data = child_growth)
```

```{r}
child_growth_df = 
  add_predictions(pwl_mod) %>% 
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = .5) +
  geom_line(aes(y = pred), color = "red")
```

Use CV to compare models
```{r}
cv_df = 
  crossv_mc(child_growth, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

Fit models and extract RMSE
```{r}
cv_df = 
  cv_df %>% 
  mutate(
    linear_mod  = map(train, ~lm(armc ~ weight, data = .x)),
    pwl_mod     = map(train, ~lm(armc ~ weight + weight_cp, data = .x)),
    smooth_mod  = map(train, ~gam(armc ~ s(weight), data = as_tibble(.x)))) %>% 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, ~rmse(model = .x, data = .y)),
    rmse_pwl    = map2_dbl(pwl_mod, test, ~rmse(model = .x, data = .y)),
    rmse_smooth = map2_dbl(smooth_mod, test, ~rmse(model = .x, data = .y)))
```

Look at RMSE distributions
```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    rmse_linear:rmse_smooth,
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin() # OR use boxplot()
```




